---
title: "Final Project"
author: "Emely Gazarov, Ryo Iwata, Maria Ramirez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(knitr)
library(car)
# to control default number formatting, no need for specifying digits in kable.
options(scipen = 3, digits = 5) 
```

```{r}
all_addiction_data = read_rds("./addiction.rds")
```

```{r}
all_addiction_data
```

```{r}
all_addiction_data |>
  filter(SUD.is_Alchohol == 1) |>
  filter(SUD.is_sum >= 2) |>
  filter(SUD.is_Cocaine == 1)


```

# Things that can predict
- number of sober days
- days since baseline
- age
- gender
- education
- drop out yes no
- SUD is alcholoc
- SUD is other
- Social support
  - family
  - friends
  - sig other
  - total
- Substance use history
  - Tried tobacco/alcohol
  - Age of first use
  - Regular use
  - Age of regular use
- AA/NA affiliation
  - Lifetime number of meetings
  - Last year number of meetings
  - Degree of affiliation
  - Positive thoughts
  - Negative thoughts
- Stressful life
  - happened
  - witnessed
  - learned about
  - exposed
  - total of all things
- Childhood
- Religion
  - Religious affiliation
  - Positive
  - Negative
- Life quality

# Things we can predict
- impression change
- length of stay
- Life quality
- Commitment to change
- Cravings
  - baseline
  - baseline vs followup
- Impression of change
