---
title: "GMS6025C: Final Project 1 Part 3 Modeling"
author: "Emely Gazarov, Ryo Iwata, Maria Ramirez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(knitr)
library(car)
# install.packages("ggalluvial")
library(ggalluvial)
# install.packages("circlize")
library(circlize)
library(stringr)
# install.packages("cowplot")
library(cowplot)
# install.packages("ggpie")
library(ggpie)
# install.packages("ggheatmap")
library(ggheatmap)
# install.packages("ggcorrplot")
library(ggcorrplot)

options(scipen = 3, digits = 5) 
```

# About the data

**Name:** Addiction Patient Assessment Measures

**Description:** Data are gathered from a residential treatment facility for substance use disorders. Data include item-level responses for all questionnaires listed below. Data are available at initiation of treatment, after one month of treatment, and at treatment discharge.

# Overview of variables in the data

- Variables that end in `.BL`, `.FU`, `.DC` are indicating that the score was collected at baseline before treatment, 30 days following up after baseline, and during discharge respectively.

## Independent variables

### Basic Demographics

- age
    - `age`
- gender
    - `gender`
- education
    - `education`
- number of days sober
    - `sober_days`
- If a patient dropped out of the program or not
    - `dropout_yn`

### Substance use disorders
  
- Flag of whether or not the patient is being treated for substance in name
    - `SUD_is_Alchohol`
    - `SUD_is_Opioid`
    - `SUD_is_Cannabis`
    - `SUD_is_depressants_anxiolytic`
    - `SUD_is_Cocaine`
    - `SUD_is_Other_stimulant`
    - `SUD_is_Hallucinogen`
    - `SUD_is_Nicotine`
    - `SUD_is_Inhalant`
    - `SUD_is_psychoactive`         
  
- Number of substances being treated for
  - Total number of all substances that patient is being treated for
    - `SUD.sum`
  - Total number of legal substances that patient is being treated for. (SUD_is_Alchohol, SUD_is_Cannabis, SUD_is_Nicotine, SUD_is_Inhalant)
    - `SUD.sum_legal`
  - Total number of legal substances that patient is being treated for. (SUD_is_Opioid, SUD_is_depressants_anxiolytic, SUD_is_Cocaine, SUD_is_Other_stimulant, SUD_is_psychoactive, SUD_is_Hallucinogen)
    - `SUD.sum_illegal`

- Flag of whether or not the patient is being treated for legal or illegals substances
    - If patient is being treated for any of the following: SUD_is_Alchohol, SUD_is_Cannabis, SUD_is_Nicotine, SUD_is_Inhalant
      - `SUD.uses_legal`
    - If patient is being treated for any of the following: SUD_is_Opioid, SUD_is_depressants_anxiolytic, SUD_is_Cocaine, SUD_is_Other_stimulant, SUD_is_psychoactive, SUD_is_Hallucinogen
      - `SUD.uses_illegal`  
  
### Beliefs about AA/NA (TSPEQ)
- Categories based on number of AA/NA meetings attended throughout life
  - `aana_life`
- Categories based on number of AA/NA meetings attended last year
  - `aana_pastyear`
- Total score of questions gauging positive opinions of AA/NA
  - `aana_positive`
- Total score of questions gauging negative opinions of AA/NA
  - `aana_negative`
    
### Childhood Experiences (ACE)
- Binary flags of whether the specified type of negative experience happened during childhood of patient
  - `childhood.verbal_abuse`
  - `childhood.physical_abuse`
  - `childhood.sexual_abuse`
  - `childhood.alone`
  - `childhood.neglected`
  - `childhood.divorced`
  - `childhood.parent_was_abused`
  - `childhood.other_was_addicted`
  - `childhood.other_was_stressed`
  - `childhood.other_was_prisoned`
- Sum of number of types of negative experiences that happened during patient's childhood
  - `childhood_sum`

### Social Support (MSPSS)
- Total score for questions that measure the perceived level of social support the respondent receives from each of the three subcategories:
  - family
    - `social.family`
  - friends
    - `social.friends`
  - significant other
    - `social.sig_other`
  - Sum of family, friends, and significant other
    - `social`

### Spiritual Experiences (Brief R-COPE)
- Total score of questions asking the frequency of spiritual coping experience, ranging from 0) Not at all to 3) Nearly every day.
  - Positive coping
    - `religion_pos`
  - Negative coping
    - `religion_neg`
- Religion
  - Binary flag of if a patient practices religion that is in the name of the variable
    - `rel.is_Atheist`
    - `rel.is_Agnostic`
    - `rel.is_Protestant`
    - `rel.is_Catholic`
    - `rel.is_Muslim`
    - `rel.is_Jewish`
    - `rel.is_Hindu`
    - `rel.is_Buddhist`
    - `rel.is_Baptist`
    - `rel.is_No_affiliation`
    - `rel.is_Non_denominational_Christian`
    - `rel.is_Other`
  - Binary flag of if patient is religious or not
    - `rel.is_religious`
    - `rel.is_not_religious`

### Stressful Life Experiences (LEC-5)
  - Binary flags of whether a listed stressful life event was either 1) experienced the life event personally, 2) witnessed it happen to someone else, 3) learned about it happening from a close friend or family member, 4) experienced the event due to their job.
    - Natural disaster
      - `stress.natural_disaster.to_subj`
      - `stress.natural_disaster.witnessed`
      - `stress.natural_disaster.learned`
      - `stress.natural_disaster.exposed`
    - Fire or explosion
      - `stress.fire.to_subj`
      - `stress.fire.witnessed`
      - `stress.fire.learned`
      - `stress.fire.exposed`
    - Transportation accident
      - `stress.transportation_accident.to_subj`
      - `stress.transportation_accident.witnessed`
      - `stress.transportation_accident.learned`
      - `stress.transportation_accident.exposed`
    - Non-transportation accident
      - `stress.serious_accident.to_subj`
      - `stress.serious_accident.witnessed`
      - `stress.serious_accident.learned`
      - `stress.serious_accident.exposed`
    - Exposure to toxic substance
      - `stress.toxic.to_subj`
      - `stress.toxic.witnessed`
      - `stress.toxic.learned`
      - `stress.toxic.exposed`
    - Physically assaulted
      - `stress.physical_assault.to_subj`
      - `stress.physical_assault.witnessed`
      - `stress.physical_assault.learned`
      - `stress.physical_assault.exposed`
    - Assaulted with a weapon
      - `stress.weapon_assault.to_subj`
      - `stress.weapon_assault.witnessed`
      - `stress.weapon_assault.learned`
      - `stress.weapon_assault.exposed`
    - Sexual assault
      - `stress.sexual_assault.to_subj`
      - `stress.sexual_assault.witnessed`
      - `stress.sexual_assault.learned`
      - `stress.sexual_assault.exposed`
    - Unwanted or uncomfortable sexual experience.
      - `stress.unwanted_sexual.to_subj`
      - `stress.unwanted_sexual.witnessed`
      - `stress.unwanted_sexual.learned`
      - `stress.unwanted_sexual.exposed`
    - Combat or exposure to war-zone 
      - `stress.combat.to_subj`
      - `stress.combat.witnessed`
      - `stress.combat.learned`
      - `stress.combat.exposed`
    - Captivity 
      - `stress.captivity.to_subj`
      - `stress.captivity.witnessed`
      - `stress.captivity.learned`
      - `stress.captivity.exposed`
    - Life-threatening illness or injury.
      - `stress.illness.to_subj`
      - `stress.illness.witnessed`
      - `stress.illness.learned`
      - `stress.illness.exposed`
    - Severe human suffering.
      - `stress.severe_suffering.to_subj`
      - `stress.severe_suffering.witnessed`
      - `stress.severe_suffering.learned`
      - `stress.severe_suffering.exposed`
    - Sudden violent death (for example, homicide, suicide).
      - `stress.sudden_violent_death.to_subj`
      - `stress.sudden_violent_death.witnessed`
      - `stress.sudden_violent_death.learned`
      - `stress.sudden_violent_death.exposed`
    - Sudden accidental death.
      - `stress.sudden_accidental_death.to_subj`
      - `stress.sudden_accidental_death.witnessed`
      - `stress.sudden_accidental_death.learned`
      - `stress.sudden_accidental_death.exposed`
    - Serious injury, harm, or death you caused to someone else.
      - `stress.harm_to_others.to_subj`
      - `stress.harm_to_others.witnessed`
      - `stress.harm_to_others.learned`
      - `stress.harm_to_others.exposed`
    - Any other very stressful event or experience.
      - `stress.other.to_subj`
      - `stress.other.witnessed`
      - `stress.other.learned`
      - `stress.other.exposed`
    - Total score of number of catgories that happened to the patient personally and/or witnessed happend to someone else 
      - `stress_to_subj`
      - `stress_to_subj_and_wit`
      - `stress_wit`

### Substance use history
- Binary flag of whether or not patient has used the following:
  - Tobacco
    - `history.tobacco`
  - Alcohol
    - `history.alcohol`
  - Other
    - `history.other`

## Dependent variables

### Length of stay in treatment
- number of days in treatment
  - `treatment_days`
  
### Alcohol/Drug Craving (PACS)
  - Total score that assesses the alcohol/drug cravings the patient has experienced within the past week based on a variety of craving characteristics.
    - `craving.BL` 
    - `craving.FU`
    - `craving.DC`
  
### Quality of Life (WHOQOL-BREF)
- Total scores of assessment that measures different aspects of quality of life
  - general
    - `qol_general_health.BL`
    - `qol_general_health.FU`
    - `qol_general_health.DC`
  - physical
    - `qol_physical.BL`
    - `qol_physical.FU`
    - `qol_physical.DC`
  - psychological health
    - `qol_psych.BL`
    - `qol_psych.FU`
    - `qol_psych.DC`
  - social relationships
    - `qol_social.BL`
    - `qol_social.FU`
    - `qol_social.DC`
  - environment
    - `qol_env.BL`
    - `qol_env.FU`
    - `qol_env.DC`
  - Total scores
    - `qol.BL`
    - `qol.FU`
    - `qol.DC`
  - Recalculating quality of life scores without social component
    - `no_social_qol.BL`

### Commitment to Sobriety (CSS-5)
- Total score of questions gauging patient’s commitment to abstinence from alcohol and drug use
  - `commit.BL` 
  - `commit.DC` 
  - `commit.FU` 
- The difference of commitment to sobriety score between discharge and baseline
  - `commit_change`

### Alcohol/Drug Abstinence Experiences (DAASE)
- Total score of questions that assesses patient’s self-confidence in their ability to remain abstinent from alcohol in various situations.
  - Negative Affect
    - `abstain_neg.BL`
    - `abstain_neg.FU`
    - `abstain_neg.DC`
  - Social/Positive
    - `abstain_pos.BL`
    - `abstain_pos.FU`
    - `abstain_pos.DC`
  - Physical ailments
    - `abstain_phy.BL`
    - `abstain_phy.FU`
    - `abstain_phy.DC` 
  - Situations associated with cravings
    - `abstain_crv.BL`
    - `abstain_crv.FU`
    - `abstain_crv.DC` 
  - All previously mentioned conditions
    - `abstain_total.BL`
    - `abstain_total.FU`
    - `abstain_total.DC` 

```{r}
cleaned_addiction_data = read_rds("./addiction.rds")
```


# First hypothesis: 

```{r}
commitment_model <- lm(commit_change ~ age 
                + gender
                + education
                + rel.is_religious
                + religion_pos
                + aana_past_year
                + SUD.sum_illegal
                + SUD.sum_legal
                ,cleaned_addiction_data)
```


$$
\begin{split}
\hat{y} = \beta_0 & + \beta_{Age}Age \\
& + \beta_{Male}Male? \\
& + \beta_{EduAA}EduAA? + \beta_{EduBS}EduBS? + \beta_{EduMS}EduMS? + \beta_{EduPhD}EduPhD? \\
& + \beta_{Religious}Religious? + \beta_{PositiveReligion}PositiveReligion \\
& + \beta_{AANAParticipation}AANAParticipation \\
& + \beta_{NumIllegal}NumIllegal + \beta_{NumIllegal}NumIllegal
\end{split}
$$

$$
\begin{split}
P(A \cup B) &= P(A) + P(B) - P(A \cap B) \\
&= 0.5 + 0.2 - 0.3 \\
&= 0.4
\end{split}
]
$$

```{r}
summary(commitment_model)
```
- Hypothesis: Individuals with higher involvement in AA/NA activities will show a greater increase in their CSS-5 relative change scores.

### Looking to see if difference of commitment to sobriety from baseline to discharge is normally distributed

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=commit_change)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of difference of commitment to sobriety at baseline to discharge",
    y = "Commitment to change quantiles",
    x = "Normal distribution quantiles")
```

### Looking to see if difference of commitment to sobriety from baseline to discharge is normally distributed

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(commit.BL)) +
  geom_histogram() +
  labs(
    title = "Distribution of baseline abstaining confidence")
```

### Looking to see if difference of commitment to sobriety from baseline to discharge is normally distributed

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=commit.BL)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of commitment to change at baseline",
    y = "Commitment to change quantiles",
    x = "Normal distribution quantiles")
```

# Looking for other normally distributed things to predict

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=craving.BL)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of craving scale at baseline",
    y = "Craving scale quantiles",
    x = "Normal distribution quantiles")
```

##

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=abstain_total.BL)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of abstaining confidence at baseline",
    y = "Abstaining confidence quantiles",
    x = "Normal distribution quantiles")
```

##

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=qol.BL)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of quality of life at baseline",
    y = "Quality of life quantiles",
    x = "Normal distribution quantiles")
```

##

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(qol.BL)) +
  geom_histogram() +
  labs(
    title = "Distribution of quality of life")
```

##

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=no_social_qol.BL)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of quality of life without social component at baseline",
    y = "Quality of life quantiles",
    x = "Normal distribution quantiles")
```

##

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(no_social_qol.BL)) +
  geom_histogram() +
  labs(
    title = "Distribution of quality of life without social component")
```

# Second hypothesis

```{r}
qol_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + rel.is_religious
                + religion_pos
                + aana_past_year
                + aana_positive
                + social
                + stress_to_subj
                ,cleaned_addiction_data)
```

##

```{r}
summary(qol_model)
```

##

```{r}

Anova(qol_model, type=2, digits=4) |>
tidy() |>
  kable() 
```

\newpage
# Plotting the adjusted data

```{r}
tidy_qol_model <- tidy(qol_model)

# Getting the slope and intercept for perceived social support
coef_social <- 
  pull(filter(tidy_qol_model, 
              term == "social"), 
       estimate)

intercept <- pull(filter(tidy_qol_model, term == "(Intercept)"), estimate)

# Getting the slope of the other variables

stress_to_subj_slope <- pull(filter(tidy_qol_model, term == "stress_to_subj"), estimate)
aana_positive_slope <- pull(filter(tidy_qol_model, term == "aana_positive"), estimate)
religion_pos_slope <- pull(filter(tidy_qol_model, term == "religion_pos"), estimate)

# Getting the mean of other variables

stress_to_subj_mean <- cleaned_addiction_data |>
  summarize(Mean = mean(stress_to_subj)) |>
  pull(Mean)

aana_positive_mean <- cleaned_addiction_data |>
  summarize(Mean = mean(aana_positive)) |>
  pull(Mean)

religion_pos_mean <- cleaned_addiction_data |>
  summarize(Mean = mean(religion_pos)) |>
  pull(Mean)

```

- Based on the following formula:
  
- We are adjusting for other things. 

##

```{r}
cleaned_addiction_data |>
  ggplot(aes(social, no_social_qol.BL, color=education)) +
  geom_point() +
  geom_abline(slope=coef_social, intercept = intercept + 
                stress_to_subj_slope * stress_to_subj_mean +
                aana_positive_slope * aana_positive_mean +
                religion_pos_slope * religion_pos_mean, linewidth=1) +
  facet_wrap(vars(education)) +
  labs(title="Quality of life adjusted for stress, AANA positive, positive religious coping")
```


##

```{r}
cleaned_addiction_data |>
  ggplot(aes(social, no_social_qol.BL, color=rel.is_religious)) +
  geom_point() +
  geom_abline(slope=coef_social, intercept = intercept + 
                stress_to_subj_slope * stress_to_subj_mean +
                aana_positive_slope * aana_positive_mean +
                religion_pos_slope * religion_pos_mean, linewidth=1) +
  facet_wrap(vars(rel.is_religious)) +
  labs(title="Quality of life adjusted for stress, AANA positive, positive religious coping")
```

```{r}
cleaned_addiction_data |>
  ggplot(aes(social, no_social_qol.BL, color=gender)) +
  geom_point() +
  geom_abline(slope=coef_social, intercept = intercept + 
                stress_to_subj_slope * stress_to_subj_mean +
                aana_positive_slope * aana_positive_mean +
                religion_pos_slope * religion_pos_mean, linewidth=1) +
  facet_wrap(vars(gender)) +
  labs(title="Quality of life adjusted for stress, AANA positive, positive religious coping")
```

\newpage
# Looking at Model Diagnostics 

```{r}
interaction_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + rel.is_religious * religion_pos
                + aana_past_year
                + aana_positive
                + social
                + stress_to_subj
                ,cleaned_addiction_data)

social_only_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + social
                ,cleaned_addiction_data)

with_religion_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + social
                + rel.is_religious * religion_pos
                ,cleaned_addiction_data)

with_aana_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + social
                + aana_past_year
                + aana_positive
                ,cleaned_addiction_data)

with_stress_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + social
                + stress_to_subj
                ,cleaned_addiction_data)
```

\newpage

## Akaike Information Criterion (AIC) Scores of models predicting quality of life with perceived level of social support

```{r}
kable(AIC(qol_model, 
          interaction_model, 
          social_only_model, 
          with_aana_model, 
          with_religion_model, 
          with_stress_model))
```

- **Original model that takes into account perceived level of social support with degree of positive religious coping, AA/NA positive regard, and number of types of stressful life circumstances has lowest AIC score.**
  - `qol_model`
  
- Other models were: 
  - based on original model and included interactions terms between if a patient is religious or not with degree of positive religious coping (`interaction_model`)
  - only considering perceived level of social support (`perceived level of social support`)
  - additionally either only considered:
    - degree of positive religious coping (`with_religion_model`)
    - AA/NA positive regard (`with_aana_model`)
    - number of types of stressful life circumstances (`with_stress_model`)**
