---
title: "GMS6025C: Final Project 1 Part 2 Exploratory Data Analysis"
author: "Emely Gazarov, Ryo Iwata, Maria Ramirez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(knitr)
library(car)
# install.packages("ggalluvial")
library(ggalluvial)
# install.packages("circlize")
library(circlize)
library(stringr)
# install.packages("cowplot")
library(cowplot)
# install.packages("ggpie")
library(ggpie)
# to control default number formatting, no need for specifying digits in kable.
# install.packages("ggheatmap")
library(ggheatmap)
# install.packages("ggcorrplot")
library(ggcorrplot)

options(scipen = 3, digits = 5) 
```

# About the data

**Name:** Addiction Patient Assessment Measures

**Description:** Data are gathered from a residential treatment facility for substance use disorders. Data include item-level responses for all questionnaires listed below. Data are available at initiation of treatment, after one month of treatment, and at treatment discharge.

# Overview of available datasets

## Independent variables

### Basic Demographics

- age
  - `age`
- gender
  - `gender`
- education
  - `education`
- number of days sober
  - `sober_days`
- number of days in treatment
  - `treatment_days`
- If a subject dropped out of the program or not
  - `dropout_yn`
  
### Substance use disorders
  
- Flag of whether or not the subject is being treated for substance in name
 - `SUD_is_Alchohol`
 - `SUD_is_Opioid`
 - `SUD_is_Cannabis`
 - `SUD_is_depressants_anxiolytic`
 - `SUD_is_Cocaine`
 - `SUD_is_Other_stimulant`
 - `SUD_is_Hallucinogen`
 - `SUD_is_Nicotine`
 - `SUD_is_Inhalant`
 - `SUD_is_psychoactive`         
  
- Number of substances being treated for
  - `SUD.sum`    
    - Total number of all substances that subject is being treated for
  - `SUD.sum_legal`
    - Total number of legal substances that subject is being treated for. (SUD_is_Alchohol, SUD_is_Cannabis, SUD_is_Nicotine, SUD_is_Inhalant)
  - `SUD.sum_illegal`
    - Total number of legal substances that subject is being treated for. (SUD_is_Opioid, SUD_is_depressants_anxiolytic, SUD_is_Cocaine, SUD_is_Other_stimulant, SUD_is_psychoactive, SUD_is_Hallucinogen)

- Flag of whether or not the subject is being treated for legal or illegals substances
  - `SUD.uses_legal`
    - If subject is being treated for any of the following: SUD_is_Alchohol, SUD_is_Cannabis, SUD_is_Nicotine, SUD_is_Inhalant
  - `SUD.uses_illegal`  
    - If subject is being treated for any of the following: SUD_is_Opioid, SUD_is_depressants_anxiolytic, SUD_is_Cocaine, SUD_is_Other_stimulant, SUD_is_psychoactive, SUD_is_Hallucinogen
  
### Beliefs about AA/NA (TSPEQ)
- Categories based on number of AA/NA meetings attended throughout life
  - `aana_life`
- Categories based on number of AA/NA meetings attended last year
  - `aana_pastyear`
- Total score of questions gauging positive opinions of AA/NA
  - `aana_positive`
- Total score of questions gauging negative opinions of AA/NA
  - `aana_negative`
    
### Childhood Experiences (ACE)
- Binary flags of whether the specified type of negative experience happened during childhood of subject
  - `childhood.verbal_abuse`
  - `childhood.physical_abuse`
  - `childhood.sexual_abuse`
  - `childhood.alone`
  - `childhood.neglected`
  - `childhood.divorced`
  - `childhood.parent_was_abused`
  - `childhood.other_was_addicted`
  - `childhood.other_was_stressed`
  - `childhood.other_was_prisoned`
- Sum of number of types of negative experiences that happened during subject's childhood
  - `childhood_sum`
  
### Commitment to Sobriety (CSS-5)
- Total score of questions gauging patient’s commitment to abstinence from alcohol and drug use
  - `commit.BL` (baseline)
  - `commit.DC` (discharge)
  - `commit.FU` (followup)

### Alcohol/Drug Abstinence Experiences (DAASE)
- Total score of questions that assesses patient’s self-confidence in their ability to remain abstinent from alcohol in various situations.
  - Negative Affect
    - `abstain_neg.BL`
    - `abstain_neg.FU`
    - `abstain_neg.DC`
  - Social/Positive
    - `abstain_pos.BL`
    - `abstain_pos.FU`
    - `abstain_pos.DC`
  - Physical ailments
    - `abstain_phy.BL`
    - `abstain_phy.FU`
    - `abstain_phy.DC` 
  - Situations associated with cravings
    - `abstain_crv.BL`
    - `abstain_crv.FU`
    - `abstain_crv.DC` 
  - All previously mentioned conditions
    - `abstain_total.BL`
    - `abstain_total.FU`
    - `abstain_total.DC` 



- Social support
  - family
  - friends
  - sig other
  - total
- Substance use history
  - Tried tobacco/alcohol
  - Age of first use
  - Regular use
  - Age of regular use

- Stressful life
  - happened
  - witnessed
  - learned about
  - exposed
  - total of all things
- Religion
  - Religious affiliation
  - Positive
  - Negative
- Life quality

## Dependent variables

### Quality of Life (WHOQOL-BREF)


- impression change
- length of stay
- Life quality
- Commitment to change
- Cravings

```{r}
cleaned_addiction_data = read_rds("./addiction.rds")
```

```{r}
colnames(select(cleaned_addiction_data, starts_with("commit")))
```


# Looking at the basic demographics


## Education distribution

```{r}
cleaned_addiction_data |> ggplot(aes(fill=education, x=gender)) +
  geom_bar() +
  labs(title = "Subject education and gender count")

```

## Age distribution

```{r}
cleaned_addiction_data |>
  ggplot(aes(gender, age, color=gender)) +
  geom_violin() + 
  geom_jitter(height = 0, width = 0.3) +
  stat_summary(fun = "mean",
             geom = "crossbar",
             color = "red") +
  # geom_boxplot(width=0.1) +
  labs(title = "Age distribution") 

```

## Distribution of number of addicted substances

```{r}
cleaned_addiction_data |> ggplot(aes(SUD.sum, color=gender, fill=gender)) +
  geom_histogram(bins = 8) +
  facet_wrap(vars(gender)) + 
  scale_x_continuous(n.breaks=10) +
  scale_y_continuous(n.breaks=20) +
  labs(title = "Number of addicted substances")

```

## Number of subjects addicted to each substance

```{r}

SUD_count_df <- cleaned_addiction_data |>
  pivot_longer(colnames(select(cleaned_addiction_data, starts_with("SUD_is"))), 
               names_to = "SUD", values_to = "SUD_yn") |>
  filter(SUD_yn == "1") |>
  mutate(SUD = str_replace(SUD, "SUD_is_", "")) |>
  mutate(SUD = str_replace(SUD, "_", "\n"))


SUD_count_df |> ggplot(aes(fill=gender, x=SUD)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=15)) +
  labs(title = "Subject count by substance use disorder")
  
```

## Number of subjects practicing each religion

```{r}

religion_count_df <- cleaned_addiction_data |>
  pivot_longer(colnames(select(cleaned_addiction_data, starts_with("rel.is_"))),
               names_to = "religion", values_to = "religion_yn") |>
  filter(religion_yn == "1") |>
  mutate(religion = str_replace(religion, "rel.is_", "")) |>
  mutate(religion = str_replace(religion, "_", "\n")) |>
  mutate(religion = str_replace(religion, "_", "\n")) 

religion_count_df |> ggplot(aes(fill=gender, x=religion)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=10)) +
  labs(title = "Subject count by religion")

  
```

# Looking at correlations
```{r}

# TODO STARTING FROM HERE

cleaned_addiction_data |>
  ggplot(aes(commit.BL, commit.DC)) +
  geom_count() +
  scale_x_continuous(n.breaks=29) +
  scale_y_continuous(n.breaks=29) +
  geom_abline(slope=1, intercept = 0, color="red", size=1, alpha=3) +
  labs(title = "Commitment to sobriety baseline to discharge")


```

```{r}
cleaned_addiction_data |>
  ggplot(aes(craving.BL, craving.FU, color=SUD.uses_illegal)) +
  geom_count() +
  scale_x_continuous(n.breaks=20) +
  scale_y_continuous(n.breaks=20) +
  geom_abline(slope=1, intercept = 0, color="red", size=1, alpha=3) +
  labs(title = "Alcholol/Drug craving baseline to discharge") +
  facet_wrap(vars(SUD.uses_illegal))
  
```
```{r}
cleaned_addiction_data |>
  ggplot(aes(SUD.uses_illegal, craving.BL, color=SUD.uses_illegal)) +
  geom_violin() + 
  geom_jitter(height = 0, width = 0.3) +
  # geom_count() +
  stat_summary(fun = "mean",
             geom = "crossbar",
             color = "red") +
  # geom_boxplot(width=0.1) +
  labs(title = "Craving distribution") 

```

```{r}
cleaned_addiction_data |> 
  ggplot(aes(SUD.uses_illegal, treatment_days, color=SUD.uses_illegal)) +
  geom_violin() +
  # geom_jitter(height = 0, width = 0.1) +
  geom_count() +
  stat_summary(fun = "mean",
             geom = "crossbar",
             color = "red") +
  # geom_boxplot(width=0.1) +
  labs(title = "Craving distribution") 
  # facet_wrap(vars(SUD.uses_illegal))
```
```{r}
cleaned_addiction_data |>
  ggplot(aes(abstain_total.BL, abstain_total.DC, color=SUD.uses_illegal)) +
  geom_count() +
  # scale_x_continuous(n.breaks=20) +
  # scale_y_continuous(n.breaks=20) +
  geom_abline(slope=1, intercept = 0, color="red", size=1, alpha=3) +
  labs(title = "Abstain confidence baseline to discharge") +
  facet_wrap(vars(SUD.uses_illegal))
  
```
```{r}
cleaned_addiction_data |>
  # select(ends_with("BL")) |> 
  select(age, treatment_days, SUD.sum_illegal, SUD.sum_legal, aana_positive, aana_negative, childhood_sum, commit.BL, abstain_total.BL, qol.BL, social, religion_pos, religion_neg, stress_to_subj, stress_wit, craving.BL) |>

  cor() |>
  ggcorrplot(colors = c("red", "white", "blue"), 
             type = "upper",
             ggtheme = ggplot2::theme_gray,
             # hc.order = TRUE
             ) +
  
  labs(title = "Correlation matrix of geneal features")
  
```
```{r}
cleaned_addiction_data |>
  ggplot(aes(aana_positive, abstain_total.BL, color=SUD.uses_illegal)) +
  geom_count() +
  # scale_x_continuous(n.breaks=20) +
  # scale_y_continuous(n.breaks=20) +
  labs(title = "AA/NA positive regard and abstain confidence") +
  facet_wrap(vars(SUD.uses_illegal))
  
```

```{r}
cleaned_addiction_data |>
  ggplot(aes(qol.BL, abstain_total.BL, color=SUD.uses_illegal)) +
  geom_count() +
  # scale_x_continuous(n.breaks=20) +
  # scale_y_continuous(n.breaks=20) +
  labs(title = "Quality of life and abstain confidence") +
  facet_wrap(vars(SUD.uses_illegal))
  
```

```{r}
cleaned_addiction_data |>
  ggplot(aes(social, qol.BL, color=SUD.uses_illegal)) +
  # geom_count() +
  geom_jitter() +
  # scale_x_continuous(n.breaks=20) +
  # scale_y_continuous(n.breaks=20) +
  labs(title = "Abstain confidence baseline to discharge") +
  facet_wrap(vars(SUD.uses_illegal))
  
```

```{r}
cleaned_addiction_data |>
  ggplot(aes(craving.BL, abstain_total.BL, color=SUD.uses_illegal)) +
  geom_count() +
  # scale_x_continuous(n.breaks=20) +
  # scale_y_continuous(n.breaks=20) +
  labs(title = "Craving to abstain confidence") +
  facet_wrap(vars(SUD.uses_illegal))
  
```
