---
title: "Final Project"
author: "Emely Gazarov, Ryo Iwata, Maria Ramirez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(knitr)
library(car)
# install.packages("ggalluvial")
library(ggalluvial)
# install.packages("circlize")
library(circlize)
library(stringr)
# install.packages("cowplot")
library(cowplot)
# install.packages("ggpie")
library(ggpie)
# to control default number formatting, no need for specifying digits in kable.
# install.packages("ggheatmap")
library(ggheatmap)
# install.packages("ggcorrplot")
library(ggcorrplot)

options(scipen = 3, digits = 5) 
```

```{r}
cleaned_addiction_data = read_rds("./addiction.rds") |>
  # mutate(commit_change = (commit.DC - commit.BL) / (commit.BL + commit.DC))
  mutate(commit_change = (commit.DC - commit.BL))

```

```{r}
cleaned_addiction_data <- cleaned_addiction_data |>
  mutate(no_social_qol.BL = (qol_general_health.BL + qol_psych.BL + qol_physical.BL + qol_env.BL)/23)
```

```{r}
cleaned_addiction_data 
```

# Overview of what we're working with

### Things that can predict
- number of sober days
- days since baseline
- age
- gender
- education
- drop out yes no
- SUD is alcholoc
- SUD is other
- Social support
  - family
  - friends
  - sig other
  - total
- Substance use history
  - Tried tobacco/alcohol
  - Age of first use
  - Regular use
  - Age of regular use
- AA/NA affiliation
  - Lifetime number of meetings
  - Last year number of meetings
  - Degree of affiliation
  - Positive thoughts
  - Negative thoughts
- Stressful life
  - happened
  - witnessed
  - learned about
  - exposed
  - total of all things
- Childhood
- Religion
  - Religious affiliation
  - Positive
  - Negative
- Life quality

### Things we can predict
- impression change
- length of stay
- Life quality
- Commitment to change
- Cravings
  - baseline
  - baseline vs followup
- Impression of change

# First hypothesis

```{r}
commitment_model <- lm(commit_change ~ age 
                + gender
                + education
                + rel.is_religious
                + religion_pos
                + aana_past_year
                + SUD.sum_illegal
                + SUD.sum_legal
                ,cleaned_addiction_data)
```

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=commit_change)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of difference of commitment to change at baseline to discharge",
    y = "Commitment to change quantiles",
    x = "Normal distribution quantiles")
```

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(commit.BL)) +
  geom_histogram() +
  labs(
    title = "Distribution of baseline abstaining confidence")
```

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=commit.BL)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of commitment to change at baseline",
    y = "Commitment to change quantiles",
    x = "Normal distribution quantiles")
```

# Looking for other normally distributed things to predict

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=craving.BL)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of craving scale at baseline",
    y = "Craving scale quantiles",
    x = "Normal distribution quantiles")
```
```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=abstain_total.BL)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of abstaining confidence at baseline",
    y = "Abstaining confidence quantiles",
    x = "Normal distribution quantiles")
```

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=qol.BL)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of quality of life at baseline",
    y = "Quality of life quantiles",
    x = "Normal distribution quantiles")
```

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(qol.BL)) +
  geom_histogram() +
  labs(
    title = "Distribution of quality of life")
```

```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(sample=no_social_qol.BL)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = "Q-Q plot of quality of life without social component at baseline",
    y = "Quality of life quantiles",
    x = "Normal distribution quantiles")
```



```{r}
# and now the Q-Q plots
cleaned_addiction_data |>
  ggplot(aes(no_social_qol.BL)) +
  geom_histogram() +
  labs(
    title = "Distribution of quality of life without social component")
```




```{r}
qol_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + rel.is_religious
                + religion_pos
                + aana_past_year
                + aana_positive
                + social
                + stress_to_subj

                ,cleaned_addiction_data)
```

```{r}
summary(qol_model)
```

```{r}

Anova(qol_model, type=2, digits=4) |>
tidy() |>
  kable() 
```

\newpage
# Looking at model metrics

```{r}
interaction_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + rel.is_religious * religion_pos
                + aana_past_year
                + aana_positive
                + social
                + stress_to_subj
                ,cleaned_addiction_data)

social_only_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + social
                ,cleaned_addiction_data)

with_religion_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + social
                + rel.is_religious * religion_pos
                ,cleaned_addiction_data)

with_aana_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + social
                + aana_past_year
                + aana_positive
                ,cleaned_addiction_data)

with_stress_model <- lm(no_social_qol.BL ~ age 
                + gender
                + education
                + SUD.sum_illegal
                + SUD.sum_legal
                + social
                + stress_to_subj
                ,cleaned_addiction_data)
```

\newpage
```{r}
kable(AIC(qol_model, interaction_model, social_only_model, with_aana_model, with_religion_model, with_stress_model))
```


```{r}

```


