---
title: "GMS6025C: Final Project 1 Part 2 Exploratory Data Analysis"
author: "Emely Gazarov, Ryo Iwata, Maria Ramirez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(knitr)
library(car)
# install.packages("ggalluvial")
library(ggalluvial)
# install.packages("circlize")
library(circlize)
library(stringr)
# install.packages("cowplot")
library(cowplot)
# install.packages("ggpie")
library(ggpie)
# to control default number formatting, no need for specifying digits in kable.
# install.packages("ggheatmap")
library(ggheatmap)
# install.packages("ggcorrplot")
library(ggcorrplot)

options(scipen = 3, digits = 5) 
```

# About the data

**Name:** Addiction Patient Assessment Measures

**Description:** Data are gathered from a residential treatment facility for substance use disorders. Data include item-level responses for all questionnaires listed below. Data are available at initiation of treatment, after one month of treatment, and at treatment discharge.

# Overview of available datasets

- Variables that end in `.BL`, `.FU`, `.DC` are indicating that the score was collected at baseline before treatment, 30 days following up after baseline, and during discharge respectively.

## Independent variables

### Basic Demographics

- age
    - `age`
- gender
    - `gender`
- education
    - `education`
- number of days sober
    - `sober_days`
- If a patient dropped out of the program or not
    - `dropout_yn`

### Substance use disorders
  
- Flag of whether or not the patient is being treated for substance in name
    - `SUD_is_Alchohol`
    - `SUD_is_Opioid`
    - `SUD_is_Cannabis`
    - `SUD_is_depressants_anxiolytic`
    - `SUD_is_Cocaine`
    - `SUD_is_Other_stimulant`
    - `SUD_is_Hallucinogen`
    - `SUD_is_Nicotine`
    - `SUD_is_Inhalant`
    - `SUD_is_psychoactive`         
  
- Number of substances being treated for
  - Total number of all substances that patient is being treated for
    - `SUD.sum`    
  - Total number of legal substances that patient is being treated for. (SUD_is_Alchohol, SUD_is_Cannabis, SUD_is_Nicotine, SUD_is_Inhalant)
    - `SUD.sum_legal`
  - Total number of legal substances that patient is being treated for. (SUD_is_Opioid, SUD_is_depressants_anxiolytic, SUD_is_Cocaine, SUD_is_Other_stimulant, SUD_is_psychoactive, SUD_is_Hallucinogen)
    - `SUD.sum_illegal`

- Flag of whether or not the patient is being treated for legal or illegals substances
    - If patient is being treated for any of the following: SUD_is_Alchohol, SUD_is_Cannabis, SUD_is_Nicotine, SUD_is_Inhalant
      - `SUD.uses_legal`
    - If patient is being treated for any of the following: SUD_is_Opioid, SUD_is_depressants_anxiolytic, SUD_is_Cocaine, SUD_is_Other_stimulant, SUD_is_psychoactive, SUD_is_Hallucinogen
      - `SUD.uses_illegal`  
  
### Beliefs about AA/NA (TSPEQ)
- Categories based on number of AA/NA meetings attended throughout life
  - `aana_life`
- Categories based on number of AA/NA meetings attended last year
  - `aana_pastyear`
- Total score of questions gauging positive opinions of AA/NA
  - `aana_positive`
- Total score of questions gauging negative opinions of AA/NA
  - `aana_negative`
    
### Childhood Experiences (ACE)
- Binary flags of whether the specified type of negative experience happened during childhood of patient
  - `childhood.verbal_abuse`
  - `childhood.physical_abuse`
  - `childhood.sexual_abuse`
  - `childhood.alone`
  - `childhood.neglected`
  - `childhood.divorced`
  - `childhood.parent_was_abused`
  - `childhood.other_was_addicted`
  - `childhood.other_was_stressed`
  - `childhood.other_was_prisoned`
- Sum of number of types of negative experiences that happened during patient's childhood
  - `childhood_sum`

### Alcohol/Drug Abstinence Experiences (DAASE)
- Total score of questions that assesses patient’s self-confidence in their ability to remain abstinent from alcohol in various situations.
  - Negative Affect
    - `abstain_neg.BL`
    - `abstain_neg.FU`
    - `abstain_neg.DC`
  - Social/Positive
    - `abstain_pos.BL`
    - `abstain_pos.FU`
    - `abstain_pos.DC`
  - Physical ailments
    - `abstain_phy.BL`
    - `abstain_phy.FU`
    - `abstain_phy.DC` 
  - Situations associated with cravings
    - `abstain_crv.BL`
    - `abstain_crv.FU`
    - `abstain_crv.DC` 
  - All previously mentioned conditions
    - `abstain_total.BL`
    - `abstain_total.FU`
    - `abstain_total.DC` 

### Social Support (MSPSS)
- Total score for questions that measure the perceived level of social support the respondent receives from each of the three subcategories:
  - family
    - `social.family`
  - friends
    - `social.friends`
  - significant other
    - `social.sig_other`
  - Sum of family, friends, and significant other
    - `social`

### Spiritual Experiences (Brief R-COPE)
- Total score of questions asking the frequency of spiritual coping experience, ranging from 0) Not at all to 3) Nearly every day.
  - Positive coping
    - `religion_pos`
  - Negative coping
    - `religion_neg`
- Religion
  - Binary flag of if a patient practices religion that is in the name of the variable
    - `rel.is_Atheist`
    - `rel.is_Agnostic`
    - `rel.is_Protestant`
    - `rel.is_Catholic`
    - `rel.is_Muslim`
    - `rel.is_Jewish`
    - `rel.is_Hindu`
    - `rel.is_Buddhist`
    - `rel.is_Baptist`
    - `rel.is_No_affiliation`
    - `rel.is_Non_denominational_Christian`
    - `rel.is_Other`
  - Binary flag of if patient is religious or not
    - `rel.is_religious`
    - `rel.is_not_religious`

### Stressful Life Experiences (LEC-5)
  - Binary flags of whether a listed stressful life event was either 1) experienced the life event personally, 2) witnessed it happen to someone else, 3) learned about it happening from a close friend or family member, 4) experienced the event due to their job.
    - Natural disaster
      - `stress.natural_disaster.to_subj`
      - `stress.natural_disaster.witnessed`
      - `stress.natural_disaster.learned`
      - `stress.natural_disaster.exposed`
    - Fire or explosion
      - `stress.fire.to_subj`
      - `stress.fire.witnessed`
      - `stress.fire.learned`
      - `stress.fire.exposed`
    - Transportation accident
      - `stress.transportation_accident.to_subj`
      - `stress.transportation_accident.witnessed`
      - `stress.transportation_accident.learned`
      - `stress.transportation_accident.exposed`
    - Non-transportation accident
      - `stress.serious_accident.to_subj`
      - `stress.serious_accident.witnessed`
      - `stress.serious_accident.learned`
      - `stress.serious_accident.exposed`
    - Exposure to toxic substance
      - `stress.toxic.to_subj`
      - `stress.toxic.witnessed`
      - `stress.toxic.learned`
      - `stress.toxic.exposed`
    - Physically assaulted
      - `stress.physical_assault.to_subj`
      - `stress.physical_assault.witnessed`
      - `stress.physical_assault.learned`
      - `stress.physical_assault.exposed`
    - Assaulted with a weapon
      - `stress.weapon_assault.to_subj`
      - `stress.weapon_assault.witnessed`
      - `stress.weapon_assault.learned`
      - `stress.weapon_assault.exposed`
    - Sexual assault
      - `stress.sexual_assault.to_subj`
      - `stress.sexual_assault.witnessed`
      - `stress.sexual_assault.learned`
      - `stress.sexual_assault.exposed`
    - Unwanted or uncomfortable sexual experience.
      - `stress.unwanted_sexual.to_subj`
      - `stress.unwanted_sexual.witnessed`
      - `stress.unwanted_sexual.learned`
      - `stress.unwanted_sexual.exposed`
    - Combat or exposure to war-zone 
      - `stress.combat.to_subj`
      - `stress.combat.witnessed`
      - `stress.combat.learned`
      - `stress.combat.exposed`
    - Captivity 
      - `stress.captivity.to_subj`
      - `stress.captivity.witnessed`
      - `stress.captivity.learned`
      - `stress.captivity.exposed`
    - Life-threatening illness or injury.
      - `stress.illness.to_subj`
      - `stress.illness.witnessed`
      - `stress.illness.learned`
      - `stress.illness.exposed`
    - Severe human suffering.
      - `stress.severe_suffering.to_subj`
      - `stress.severe_suffering.witnessed`
      - `stress.severe_suffering.learned`
      - `stress.severe_suffering.exposed`
    - Sudden violent death (for example, homicide, suicide).
      - `stress.sudden_violent_death.to_subj`
      - `stress.sudden_violent_death.witnessed`
      - `stress.sudden_violent_death.learned`
      - `stress.sudden_violent_death.exposed`
    - Sudden accidental death.
      - `stress.sudden_accidental_death.to_subj`
      - `stress.sudden_accidental_death.witnessed`
      - `stress.sudden_accidental_death.learned`
      - `stress.sudden_accidental_death.exposed`
    - Serious injury, harm, or death you caused to someone else.
      - `stress.harm_to_others.to_subj`
      - `stress.harm_to_others.witnessed`
      - `stress.harm_to_others.learned`
      - `stress.harm_to_others.exposed`
    - Any other very stressful event or experience.
      - `stress.other.to_subj`
      - `stress.other.witnessed`
      - `stress.other.learned`
      - `stress.other.exposed`
    - Total score of number of catgories that happened to the patient personally and/or witnessed happend to someone else 
      - `stress_to_subj`
      - `stress_to_subj_and_wit`
      - `stress_wit`

### Substance use history
- Binary flag of whether or not patient has used the following:
  - Tobacco
    - `history.tobacco`
  - Alcohol
    - `history.alcohol`
  - Other
    - `history.other`

## Dependent variables

### Length of stay in treatment
- number of days in treatment
  - `treatment_days`
  
### Alcohol/Drug Craving (PACS)
  - Total score that assesses the alcohol/drug cravings the patient has experienced within the past week based on a variety of craving characteristics.
    - `craving.BL` 
    - `craving.FU`
    - `craving.DC`
  
### Quality of Life (WHOQOL-BREF)
- Total scores of assessment that measures different aspects of quality of life
  - general
    - `qol_general_health.BL`
    - `qol_general_health.FU`
    - `qol_general_health.DC`
  - physical
    - `qol_physical.BL`
    - `qol_physical.FU`
    - `qol_physical.DC`
  - psychological health
    - `qol_psych.BL`
    - `qol_psych.FU`
    - `qol_psych.DC`
  - social relationships
    - `qol_social.BL`
    - `qol_social.FU`
    - `qol_social.DC`
  - environment
    - `qol_env.BL`
    - `qol_env.FU`
    - `qol_env.DC`

### Commitment to Sobriety (CSS-5)
- Total score of questions gauging patient’s commitment to abstinence from alcohol and drug use
  - `commit.BL` 
  - `commit.DC` 
  - `commit.FU` 


```{r}
cleaned_addiction_data = read_rds("./addiction.rds")
```

# Looking at the basic demographics


## Education distribution

```{r}
cleaned_addiction_data |> ggplot(aes(fill=education, x=gender)) +
  geom_bar() +
  labs(title = "patient education and gender count")

```

## Age distribution

```{r}
cleaned_addiction_data |>
  ggplot(aes(gender, age, color=gender)) +
  geom_violin() + 
  geom_jitter(height = 0, width = 0.3) +
  stat_summary(fun = "mean",
             geom = "crossbar",
             color = "red") +
  # geom_boxplot(width=0.1) +
  labs(title = "Age distribution") 

```

## Distribution of number of addicted substances

```{r}
cleaned_addiction_data |> ggplot(aes(SUD.sum, color=gender, fill=gender)) +
  geom_histogram(bins = 8) +
  facet_wrap(vars(gender)) + 
  scale_x_continuous(n.breaks=10) +
  scale_y_continuous(n.breaks=20) +
  labs(title = "Number of addicted substances")

```

## Number of patients addicted to each substance

```{r}

SUD_count_df <- cleaned_addiction_data |>
  pivot_longer(colnames(select(cleaned_addiction_data, starts_with("SUD_is"))), 
               names_to = "SUD", values_to = "SUD_yn") |>
  filter(SUD_yn == "1") |>
  mutate(SUD = str_replace(SUD, "SUD_is_", "")) |>
  mutate(SUD = str_replace(SUD, "_", "\n"))


SUD_count_df |> ggplot(aes(fill=gender, x=SUD)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=15)) +
  labs(title = "patient count by substance use disorder")
  
```

## Number of patients practicing each religion

```{r}

religion_count_df <- cleaned_addiction_data |>
  pivot_longer(colnames(select(cleaned_addiction_data, starts_with("rel.is_"))),
               names_to = "religion", values_to = "religion_yn") |>
  filter(religion_yn == "1") |>
  mutate(religion = str_replace(religion, "rel.is_", "")) |>
  mutate(religion = str_replace(religion, "_", "\n")) |>
  mutate(religion = str_replace(religion, "_", "\n")) 

religion_count_df |> ggplot(aes(fill=gender, x=religion)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=10)) +
  labs(title = "patient count by religion")

  
```

# Looking at correlations

## Correlations across continous variables

```{r}
cleaned_addiction_data |>
  # select(ends_with("BL")) |> 
  select(age, treatment_days, SUD.sum_illegal, SUD.sum_legal, aana_positive, aana_negative, childhood_sum, commit.BL, abstain_total.BL, qol.BL, social, religion_pos, religion_neg, stress_to_subj, stress_wit, craving.BL) |>

  cor() |>
  ggcorrplot(colors = c("red", "white", "blue"), 
             type = "upper",
             ggtheme = ggplot2::theme_gray,
             # hc.order = TRUE
             ) +
  
  labs(title = "Correlation matrix of geneal features")
  
```

## Looking at relationship between commitment to sobriety at baseline to discharge

```{r}
cleaned_addiction_data |>
  ggplot(aes(commit.BL, commit.DC)) +
  geom_count() +
  scale_x_continuous(n.breaks=29) +
  scale_y_continuous(n.breaks=29) +
  geom_abline(slope=1, intercept = 0, color="red", linewidth=1, alpha=3) +
  labs(title = "Commitment to sobriety baseline to discharge")


```


## Looking at relationship between craving levels to substances at baseline to discharge

```{r}
cleaned_addiction_data |>
  ggplot(aes(SUD.uses_illegal, craving.BL, color=SUD.uses_illegal)) +
  geom_violin() + 
  geom_jitter(height = 0, width = 0.3) +
  # geom_count() +
  stat_summary(fun = "mean",
             geom = "crossbar",
             color = "red") +
  # geom_boxplot(width=0.1) +
  labs(title = "Craving distribution") 

```

## Looking at distribution of craving levels for those that use and don't use illegal substances

```{r}
cleaned_addiction_data |> 
  ggplot(aes(SUD.uses_illegal, treatment_days, color=SUD.uses_illegal)) +
  geom_violin() +
  # geom_jitter(height = 0, width = 0.1) +
  geom_count() +
  stat_summary(fun = "mean",
             geom = "crossbar",
             color = "red") +
  # geom_boxplot(width=0.1) +
  labs(title = "Craving distribution") 
  # facet_wrap(vars(SUD.uses_illegal))
```

##

```{r}
cleaned_addiction_data |>
  ggplot(aes(abstain_total.BL, abstain_total.DC, color=SUD.uses_illegal)) +
  geom_count() +
  # scale_x_continuous(n.breaks=20) +
  # scale_y_continuous(n.breaks=20) +
  geom_abline(slope=1, intercept = 0, color="red", size=1, alpha=3) +
  labs(title = "Abstain confidence baseline to discharge") +
  facet_wrap(vars(SUD.uses_illegal))
  
```

##

```{r}
cleaned_addiction_data |>
  ggplot(aes(aana_positive, abstain_total.BL, color=SUD.uses_illegal)) +
  geom_count() +
  # scale_x_continuous(n.breaks=20) +
  # scale_y_continuous(n.breaks=20) +
  labs(title = "AA/NA positive regard and abstain confidence") +
  facet_wrap(vars(SUD.uses_illegal))
  
```

##

```{r}
cleaned_addiction_data |>
  ggplot(aes(qol.BL, abstain_total.BL, color=SUD.uses_illegal)) +
  geom_count() +
  # scale_x_continuous(n.breaks=20) +
  # scale_y_continuous(n.breaks=20) +
  labs(title = "Quality of life and abstain confidence") +
  facet_wrap(vars(SUD.uses_illegal))
  
```

## 

```{r}
cleaned_addiction_data |>
  ggplot(aes(social, qol.BL, color=SUD.uses_illegal)) +
  # geom_count() +
  geom_jitter() +
  # scale_x_continuous(n.breaks=20) +
  # scale_y_continuous(n.breaks=20) +
  labs(title = "Abstain confidence baseline to discharge") +
  facet_wrap(vars(SUD.uses_illegal))
  
```

## 

```{r}
cleaned_addiction_data |>
  ggplot(aes(craving.BL, abstain_total.BL, color=SUD.uses_illegal)) +
  geom_count() +
  # scale_x_continuous(n.breaks=20) +
  # scale_y_continuous(n.breaks=20) +
  labs(title = "Craving to abstain confidence") +
  facet_wrap(vars(SUD.uses_illegal))
  
```
