---
title: "Group 04 Project 1 - Data Carpentry & Visualization Basics"
author: "Stephan Quintin, Ryo Iwata, Zachary Strickland"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(nycflights13)
```

# Group 4 members:
names here Stephan Quintin, Ryo Iwata, Zachary Strickland
Note who is assigned leader: Stephan Quintin

# Group Assignment 1
Instructions - Everyone should work on their own computer on a separate copy of this, then work together to create a single very well polished version to submit for the assignment. The assigned group leader should submit the assignment for the group, but every individual should thoroughly understand every answer. In most cases, everyone should roughly be on the same question as the same time when working together simultaneously (e.g., in class or during group meetings). When asked, you (all) need to be ready to explain the thought process behind the plan, approach, revisions, and final code that was written. Now is the time to build your understanding of the language and approach to these data carpentry and visualization methods. My expectation is that every submitted assignment will be nearly perfect given how we will be approaching this series of problems both in class and out.

I have copied together all the questions in one place so everyone doesn't have to waste their time doing so, but these are directly from the reading (R4DS 2E). I am leaving the formatting plain, so refer to the book itself for the example plots and pretty formatting.

Please make it easy to grade this! Your text answers should be in the main document (i.e., not in a code chunk), and *please make them bold so they stand out*. Write code, including using comments, as if you are writing code that will be published. Build good habits now!

## 1.2.5

### 1 
How many rows are in penguins? How many columns?

**Answer: 344 rows and 8 columns**

```{r}
# Looks at the structure of penguins
str(penguins)

```

### 2
What does the bill_depth_mm variable in the penguins data frame describe? Read the help for ?penguins to find out.

**Answer: a number denoting bill depth (millimeters)**

```{r}
# ?penguins
```

### 3
Make a scatterplot of bill_depth_mm vs. bill_length_mm. That is, make a scatterplot with bill_depth_mm on the y-axis and bill_length_mm on the x-axis. Describe the relationship between these two variables.

**No meaningful visual relationship, there might be some clusters**

```{r}

ggplot(penguins, aes(bill_depth_mm, bill_length_mm)) + geom_point()
```

### 4 
What happens if you make a scatterplot of species vs. bill_depth_mm? What might be a better choice of geom?

**It creates a scatter plot for each species. geom_violin could better show the distribution**

```{r}
ggplot(penguins, aes(species, bill_depth_mm)) + geom_point()

ggplot(penguins, aes(species, bill_depth_mm)) + geom_violin()

```

### 5
Why does the following give an error and how would you fix it?

*X and Y arguments are missing in the aesthetic function. We would fix this by defining X and Y in the aesthetic function*

```{r}
# ggplot(data = penguins, aes()) +
#  geom_point()
```

### 6
What does the na.rm argument do in geom_point()? What is the default value of the argument? Create a scatterplot where you successfully use this argument set to TRUE.

*na.rm is whether or not there is a warning of missing data points. Default value is FALSE which shows a warning.

```{r}

ggplot(penguins, aes(flipper_length_mm, body_mass_g)) + geom_point(na.rm = TRUE,)

```


### 7
Add the following caption to the plot you made in the previous exercise: “Data come from the palmerpenguins package.” Hint: Take a look at the documentation for labs().

```{r}
ggplot(penguins, aes(flipper_length_mm, body_mass_g)) + geom_point(na.rm = TRUE,) + labs(caption="Data come from the palmerpenguins package")

```

### 8
Recreate the following visualization. What aesthetic should bill_depth_mm be mapped to? And should it be mapped at the global level or at the geom level?

```{r}
ggplot(penguins, aes(flipper_length_mm, body_mass_g, colour=bill_depth_mm)) + geom_point() +  geom_smooth()

```

### 9
Run this code in your head and predict what the output will look like. 

ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)
  
**Answer: It will be a scatter plot with flipper length as the x, body mass as the y, and each point is colored based on the associated island. A smoothed line that best passes through the points going from the start to end of x will be drawn.There will be one line for each island that the penguins came from. Because `se` is `FALSE`, there will be no confidence interval drawn.
  
### 9b

```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)
  

```

### 10
Will these two graphs look different? Why/why not?

** Answer: The two graphs will look the same. This is because both sections of code are achieving the same thing in different ways. The first is setting the data and mapping for the entire plot with the `ggplot` function. `geom_point() and geom_smooth()` inherit this data and mapping because they have no arguments. The second calls `ggplot` without any arguments, so it creates an empty plot. `geom_point() and geom_smooth()` both have the data and mapping as arguments that's identical to that of the first section, so the plot will be identical. **

ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point() +
  geom_smooth()

ggplot() +
  geom_point(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  ) +
  geom_smooth(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  )
  
### 10b confirm your answer:
```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point() +
  geom_smooth()

ggplot() +
  geom_point(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  ) +
  geom_smooth(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  )

```

\newpage

## 1.4.3 Exercises

### 1
Make a bar plot of species of penguins, where you assign species to the y aesthetic. How is this plot different?

** Answer: This plot is different because it only requires 1 variable as an argument. It uses the count of data points that are within a given category as the x-value.**

```{r}

ggplot(
  data = penguins,
  mapping = aes(y = species)
) +
  geom_bar()
```

### 2
How are the following two plots different? Which aesthetic, color or fill, is more useful for changing the color of bars?

    ggplot(penguins, aes(x = species)) +
      geom_bar(color = "red")

    ggplot(penguins, aes(x = species)) +
      geom_bar(fill = "red")
      
** Answer: The first section only changes the outline of the bars to red. The second changes the color of the insides of the bar. The second if more useful because the outline is small and ususally the focus of labeling for a plot.
      
```{r}

    ggplot(penguins, aes(x = species)) +
      geom_bar(color = "red")

    ggplot(penguins, aes(x = species)) +
      geom_bar(fill = "red")
```     
  

### 3 
What does the bins argument in geom_histogram() do?

** The bins argument determines the number of bins in the histogram. **


### 4
Make a histogram of the carat variable in the diamonds dataset that is available when you load the tidyverse package. Experiment with different binwidths. What binwidth reveals the most interesting patterns?

** Answer: Setting the binwidth to 0.01 shows that there are sudden jumps in count in carats that are divisible by 0.25. This may reflect carats being rounded or cut to this size. **


```{r}
 



ggplot(diamonds, aes(carat)) +
  geom_histogram(binwidth = 0.01)

```


## 1.5.5 Exercises

### 1
The mpg data frame that is bundled with the ggplot2 package contains 234 observations collected by the US Environmental Protection Agency on 38 car models. Which variables in mpg are categorical? Which variables are numerical? (Hint: Type ?mpg to read the documentation for the dataset.) How can you see this information when you run mpg?

** 

Categorical: manufacturer, model, trans, dry, cty, class
Numerical: displ, year, cyl, hwy, fl

You can see if a variable is categorical or numerical by using the `str` function on the tibble.

**


manufacturer: chr [1:234] "audi" "audi" "audi" "audi" ...
 $ model       : chr [1:234] "a4" "a4" "a4" "a4" ...
 $ displ       : num [1:234] 1.8 1.8 2 2 2.8 2.8 3.1 1.8 1.8 2 ...
 $ year        : int [1:234] 1999 1999 2008 2008 1999 1999 2008 1999 1999 2008 ...
 $ cyl         : int [1:234] 4 4 4 4 6 6 6 4 4 4 ...
 $ trans       : chr [1:234] "auto(l5)" "manual(m5)" "manual(m6)" "auto(av)" ...
 $ drv         : chr [1:234] "f" "f" "f" "f" ...
 $ cty         : int [1:234] 18 21 20 21 16 18 18 18 16 20 ...
 $ hwy         : int [1:234] 29 29 31 30 26 26 27 26 25 28 ...
 $ fl          : chr [1:234] "p" "p" "p" "p" ...
 $ class  

```{r}


str(mpg)

```

### 2
Make a scatterplot of hwy vs. displ using the mpg data frame. Next, map a third, numerical variable to color, then size, then both color and size, then shape. How do these aesthetics behave differently for categorical vs. numerical variables?

** color: For numerical there is a gradient of lighter to darker blue as the number changes. For categorical there are different colors that are associated with each category. 

size: For numerical there is a gradient of sizes as the number changes. For categorical there are different discrete sizes that are associated with each category.

**


```{r}
 
ggplot(mpg, aes(x=hwy, y=displ, colour=year)) + geom_point()

ggplot(mpg, aes(x=hwy, y=displ, size=year)) + geom_point()

ggplot(mpg, aes(x=hwy, y=displ, colour=year, size=year)) + geom_point()

ggplot(mpg, aes(x=hwy, y=displ, colour=class)) + geom_point()

ggplot(mpg, aes(x=hwy, y=displ, size=class)) + geom_point()

ggplot(mpg, aes(x=hwy, y=displ, colour=class, size=class)) + geom_point()


```

### 3
In the scatterplot of hwy vs. displ, what happens if you map a third variable to linewidth?

**Nothing, because there are no lines in the scatterplot**
```{r}
#simple scatter plot of hwy MPG (x axis) vs engine displacement volume (y axis).
ggplot(data = mpg, aes(x = hwy, y = displ)) +
  geom_point()
#added a third variable to linewidth
#simple scatter plot of hwy MPG (x axis) vs engine displacement volume (y axis).
ggplot(data = mpg, aes(x = hwy, y = displ, linewidth = cty)) +
  geom_point()
```


### 4
What happens if you map the same variable to multiple aesthetics?

**Then the same variable will be mapped to multiple aesthetics**
```{r}
#let's try it out...
ggplot(data = mpg, aes(x = hwy, y = displ, color = cty, size = cty)) +
  geom_point()

```

### 5
Make a scatterplot of bill_depth_mm vs. bill_length_mm and color the points by species. What does adding coloring by species reveal about the relationship between these two variables? What about faceting by species?

### 6
Why does the following yield two separate legends? How would you fix it to combine the two legends?

  ggplot(
    data = penguins,
    mapping = aes(
      x = bill_length_mm, y = bill_depth_mm, 
      color = species, shape = species
    )
  ) +
    geom_point() +
    labs(color = "Species")

### 7 
Create the two following stacked bar plots. Which question can you answer with the first one? Which question can you answer with the second one?
```{r}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(position = "fill")
ggplot(penguins, aes(x = species, fill = island)) +
  geom_bar(position = "fill")
```



## 2.5 Exercises

### 1
Why does this code not work?

  my_variable <- 10
  my_varıable
  #> Error in eval(expr, envir, enclos): object 'my_varıable' not found

Look carefully! (This may seem like an exercise in pointlessness, but training your brain to notice even the tiniest difference will pay off when programming.)

** Answer: my_varıable is spelled differently than the variable that was defined, so my_varıable was not defined

### 2
Tweak each of the following R commands so that they run correctly:
```{r}
# remove the comment, i.e. # symbol from the below lines by selecting and pressing control-shift-c (i.e., ctrl-C)
# libary(todyverse)
# 
# ggplot(dTA = mpg) +
#   geom_point(maping = aes(x = displ y = hwy)) +
#   geom_smooth(method = "lm)

library(tidyverse)

ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point()
 geom_smooth(method = "lm")
```

### 3
Press Option + Shift + K / Alt + Shift + K. What happens? How can you get to the same place using the menus?

** Answer: The keyboard shortcuts appear. You can go to `Help >> Keyboard Shortcuts Help` **

### 4
Let’s revisit an exercise from the Section 1.6. Run the following lines of code. Which of the two plots is saved as mpg-plot.png? Why?

** `my_bar_plot` is saved because it is speficied in the `ggsave` function under the `plot` variable. **

```{r}
my_bar_plot <- ggplot(mpg, aes(x = class)) +
  geom_bar()
my_scatter_plot <- ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point()
#ggsave(filename = "mpg-plot.png", plot = my_bar_plot)
```

## 3.2.5 Exercises

### 1
In a single pipeline for each condition, find all flights that meet the condition:
        Had an arrival delay of two or more hours
        Flew to Houston (IAH or HOU)
        Were operated by United, American, or Delta
        Departed in summer (July, August, and September)
        Arrived more than two hours late, but didn’t leave late
        Were delayed by at least an hour, but made up over 30 minutes in flight
        
        
```{r}
#for my own edification view/str 
View(flights)
str(flights)
#flights with arrival delay of >120min
flights |> 
  filter(arr_delay > 120)
# Filter flights to those with destination of IAH or HOU 
flights |> 
  filter(dest == "IAH"|dest == "HOU")
# filter to flights operated by UA, AA, & DL
flights |>
  filter(carrier == "UA" | carrier == "AA" | carrier == "DL")
# filter to months July-Sept
flights |>
  filter(month == 7| month == 8| month == 9)
# flights that arrived >120min late, but left on time
flights |>
  filter(dep_delay == 0 & arr_delay > 120)
#flights that left at least an hour late but arrived within 30 minutes of their scheduled arrival time. THis isn't perfect but its the best I can do at this time.
flights |>
  filter(dep_delay > 60 & arr_delay <30)
```

### 2
Sort flights to find the flights with longest departure delays. Find the flights that left earliest in the morning.
```{r}
# sort(flights, dep_delay)

flights %>% 
  arrange(desc(dep_delay))

filter(flights, dep_time > 800) %>% 
  arrange(dep_time)

```

### 3
Sort flights to find the fastest flights. (Hint: Try including a math calculation inside of your function.)
```{r}
flights %>%
  arrange(desc(air_time / distance))

```

### 4
Was there a flight on every day of 2013?
**Answer: When we counted all flights that occurred in month day pair there are 365 rows corresponding to data points for each day of the year.

```{r}
# Filtering for all flights that occured in 2013
# Piping all 2013 flights and looking at all the unique month day pairings
flights %>%
  count(month,day)
```

### 5
Which flights traveled the farthest distance? Which traveled the least distance?
**Answer: Least distance was flight 1632, Farthest distance was 51.**
```{r}

flights %>%
  arrange(distance)
flights %>%
  arrange(desc(distance))
```

### 6
Does it matter what order you used filter() and arrange() if you’re using both? Why/why not? Think about the results and how much work the functions would have to do. 
**Answer:The order that the functions are used does not change the output of the data because filtering and arranging are both maintained when followed by the other respective function.**
```{r}
flights %>%
filter(arr_delay>2) %>%
arrange(arr_delay)

flights %>%
  arrange(arr_delay) %>%
filter(arr_delay>2) 

```



## 3.3.5 Exercises

### 1  
Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?
**dep_time minus sched_dep_time equals dep_delay **
 
```{r}
flights %>%
  select(dep_time, sched_dep_time, dep_delay)
```
### 2
Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
flights %>%
select(dep_time, dep_delay, arr_time, arr_delay)
flights %>%
select(starts_with("dep"), starts_with("arr") )
flights %>%
  select(contains("time"),contains("delay"),-contains("sched"),-contains("air"),-contains("hour"))
flights %>%
  select(c(4,6,7,9))
```
### 3
What happens if you specify the name of the same variable multiple times in a select() call?
It ignores redundant column calls.
```{r}
flights %>%
  select(dep_time, dep_time)
```
### 4
What does the any_of() function do? Why might it be helpful in conjunction with this vector?
variables <- c("year", "month", "day", "dep_delay", "arr_delay")
** Answer: `any_of() function allows you to select for multiple items. In this case, it can be used to filter for all columns that are in the vector without raising an error if a column in the vector is missing in the tibble. **
 
```{r}
variables <- c("year", "month", "day", "dep_delay", "arr_delay")
 
 
flights %>% select(any_of(variables))
```
### 5
Does the result of running the following code surprise you? How do the select helpers deal with upper and lower case by default? How can you change that default?
**Answer=The select function defaults to ignoring case so despite being in all caps the function still calls all columns that contain "time". To change the default you change the ignore.case argument to False.**
```{r}
flights |> select(contains("TIME"))

flights |> select(contains("TIME",ignore.case=FALSE))
```
### 6 
Rename air_time to air_time_min to indicate units of measurement and move it to the beginning of the data frame.
```{r}
flights %>%
  rename(air_time_min=air_time) %>%
  relocate(air_time_min)
```
### 7
Why doesn’t the following work, and what does the error mean?
**Answer: The code doesn't work because only the column `tailnum` was selected for, and `arr_delay` was no longer part of the new tibble. So it wouldn't have been able to filter by `arr_delay`. The error shows what function caused the error which is `arrange()`, and that the error was because `arr_delay` was not in the tibble.**
```{r}
# remove comments by selecting all desired lines of code and using ctrl-C (control-shift-c) i.e. comment/uncomment line
# 
#flights |> 
#  select(tailnum) 
#   arrange(arr_delay)
# #> Error in `arrange()`:
# #> i In argument: `..1 = arr_delay`.
# #> Caused by error:
# #> ! object 'arr_delay' not found
```

## 3.5.7 Exercises

### 1
Which carrier has the worst average delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights |> group_by(carrier, dest) |> summarize(n()))
**Answer: F9 airline is the worst**
```{r}
flights %>%
  group_by(carrier) %>%
  summarise(avg = mean(dep_delay, na.rm = TRUE) + mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(avg))

```

### 2
Find the flights that are most delayed upon departure from each destination.
```{r}
flights %>% 
  group_by(dest) %>%
  slice_max(dep_delay, n = 1) %>%
  relocate(dest)
```

### 3
How do delays vary over the course of the day. Illustrate your answer with a plot.
**Answer: Delays get progressively longer throughout the day**
```{r}
flights %>%
  mutate(delay=arr_delay+dep_delay) %>%
  ggplot() + geom_point(aes(x = dep_time, y = delay))
```

### 4 
What happens if you supply a negative n to slice_min() and friends?
**answer: with a negative n, functions like slice_min() will return all rows instead of a specified number.**
```{r}
flights %>%
slice_min(dep_delay, n = 1)
flights %>%
slice_min(dep_delay, n = -1)
```

### 5
Explain what count() does in terms of the dplyr verbs you just learned. What does the sort argument to count() do?
**Answer: `count()` function counts the number of rows that have a given value in a specified column. Then it creates a new table with the counts of each group.The sort argument order the rows in decending value.** 
 
```{r}
flights %>% count(carrier, sort=TRUE)
```
### 6
Suppose we have the following tiny data frame:
```{r}
     df <- tibble(
       x = 1:5,
       y = c("a", "b", "a", "a", "b"),
       z = c("K", "K", "L", "L", "K")
    )
```

#### 6.a
Write down what you think the output will look like, then check if you were correct, and describe what group_by() does.
**Answer: It looks like the `group_by()` function will create groups based on the rows that share the same `y` value from `df`. According to the documentation, `group_by`creates a new table that is grouped where operations can performed per group.**
```{r}

 df |>
  group_by(y)
```

#### 6.b 
Write down what you think the output will look like, then check if you were correct, and describe what arrange() does. Also comment on how it’s different from the group_by() in part (a)?
**Answer: The output might be a table that is the sorted view of `df` based on the `y` value. It's different than `group_by()` because there seems to be no group functionality with the output of this code. **
 
```{r}
 
df |>
  arrange(y)
```
#### 6.c
Write down what you think the output will look like, then check if you were correct, and describe what the pipeline does.
**Answer: The output will probably be a table that has a row for each group of rows from `df` that had a given `y` value. Then the mean of `x` was calculated for all rows in a given group.The pipeline outputs all of the `df` table into `group_by(y)` function, which creates a group of rows that had a given `y` value. These groupings are outputted to the   `summarize(mean_x = mean(x))` where the mean `x` was calculated for all rows in a given group. **
```{r}
 df |>
   group_by(y) |>
   summarize(mean_x = mean(x))
```


#### 6.d
Write down what you think the output will look like, then check if you were correct, and describe what the pipeline does. Then, comment on what the message says.
**Answer: The output will probably be a table that has a row for each group of rows from `df` that had a given `y` and `z` unique pair. Then the mean of `x` was calculated for all rows in a given group.The pipeline outputs all of the `df` table into `group_by(y, z)` function, which creates a group of rows that had a given `y` and `z` unique pair. These groupings are outputted to the   `summarize(mean_x = mean(x))` where the mean `x` was calculated for all rows in a given group.**
```{r}
df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))
```

#### 6.e

Write down what you think the output will look like, then check if you were correct, and describe what the pipeline does. How is the output different from the one in part (d).
**Answer: The output will probably be similar to the output from problem 6.d with a table that has a row for each group of rows from `df` that had a given `y` and `z` unique pair. Then the mean of `x` was calculated for all rows in a given group. The pipeline outputs all of the `df` table into `group_by(y, z)` function, which creates a group of rows that had a given `y` and `z` unique pair. These groupings are outputted to the   `summarize(mean_x = mean(x))` where the mean `x` was calculated for all rows in a given group.The main difference is that the `group` functionality is removed in this output, and new tibble is created. ** 
```{r}
df |>
   group_by(y, z) |>
   summarize(mean_x = mean(x), .groups = "drop")
```

#### 6.f
Write down what you think the outputs will look like, then check if you were correct, and describe what each pipeline does. How are the outputs of the two pipelines different?
**Answer: The output will probably be similar to the output from problem 6.d with a table that has a row for each group of rows from `df` that had a given `y` and `z` unique pair. Then the mean of `x` was calculated for all rows in a given group. The pipeline outputs all of the `df` table into `group_by(y, z)` function, which creates a group of rows that had a given `y` and `z` unique pair. These groupings are outputted to the   `summarize(mean_x = mean(x))` where the mean `x` was calculated for all rows in a given group. The main difference is that the `group` functionality is removed in this output, and new tibble is created.**
```{r}
 df |>
  group_by(y, z) |>
  summarize(mean_x = mean(x))

df |>
 group_by(y, z) |>
 mutate(mean_x = mean(x))
```