---
title: "ANOVAs"
author: "Damon G. Lamb"
date: ""
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(car)
library(tidymodels)
library(Hmisc)
```

In this in-class problem set, we will work through what happens when parameters of our experiment and the underlying distribution / error, how that impacts our model (estimates, ANOVA, etc.), and visually what the impact is. 

The objectives are to develop a deeper understanding of the core features of anova and thinking about anova as regression as a piece of our foundation for a more general linear model framework. 
Practicing our coding and working with stat_XXX in ggplot will also help prepare for homework and application beyond.


# Load data (note multiple sheets)

```{r}
n500data=readxl::read_excel("./data/ANOVA_1way.xlsx", sheet = "n500")


```


# Explore Visualization

Make a single set of plots for one sheet, including options that have summary values, all the data, plots combining visual representations, etc. Start with the basics - things you have seen in the lit or textbooks, then try to use other geoms, stats, aesthetic mappings. Explore and experiment. We will discuss and share options.

```{r}
ggplot(n500data, aes(`Group Type`, Data))+
  stat_summary( 
               geom = "bar",
               fun.data = "mean_se")+
  stat_summary(color="Red", 
               geom = "pointrange",
               size = 0.15,
               fun.data = "mean_se")+
  coord_cartesian(ylim = c(15,50))
# note - ylim will cause errors, think about why (and look at documentation)
# the argument to fun.data is from Hmisc, referenced from ggplot
# fun.data is the function used to calculate summary values passed to the geometry selected
# In the case of sdl, the default multiplier is 2 (i.e., plus or minus 2 sd)
# I do not know why.
```


You've probably seen plots like that.
You shouldn't trust them!


```{r}
ggplot(n500data, aes(x = Data, color = `Group Type`)) +
  facet_wrap(vars(`Group Type`))+
  geom_histogram(aes(y = after_stat(density)),
                 bins = 30, 
                 fill=NA) +
  geom_density()
# the after_stat for 7 allows overlapping of the density and histogram in an elegant way
# since we don't care about the specific count y (from histogram)

```

```{r}
ggplot(n500data, aes(`Group Type`, Data))+
  geom_violin()+
  geom_jitter(alpha = 0.25)+
  stat_summary(color="Red", 
               geom = "crossbar",
               size = 0.1,
               fun.data = "mean_se")+
  stat_summary(color="Blue", 
               size=0.1, 
               fun.data = "mean_sdl", 
               fun.args = list(mult = 1))

summarise(n500data, 
          .by=`Group Type`, 
          avg = mean(Data), 
          std = sd(Data))
#note - with such a high sample size, SEM or SE is sigma / sqrt(n)
```


# Choose "best"
Make a plot that most accurately (within reason) shows the data, distribution, and the model.
Think about what the model is and thus what you should include.


```{r}
# this fits a model with one factor predictor (Group Type) and with one dependent variable (Data)
# formula = 
n500mod = lm(formula = Data~`Group Type`, data = n500data) 
# Complement to Applied Regression has the valid Anova test to use. We will dig into this later.
car::Anova(n500mod)

# these tidymodel functions provide useful tibbles to look at the model, fit, etc.
glance(n500mod)

tidy(n500mod)

augment(n500mod)
```

# Model

Manually write out your model () You can refer back to our prior examples, but it is easiest to use 

$$y = \beta_0 + \beta_{B}B? + \beta_{C}C? + \beta_{D}D? + \beta_{E}E? + \epsilon$$ 
$$y = \beta_0 + \beta_{B} *0+ \beta_{C}*0 + \beta_{D}*0 + \beta_{E}*0 + \epsilon$$ 
$$y = \beta_0 + \epsilon$$ 
$$y = \beta_0 + \beta_{B} *1+ \beta_{C}*0 + \beta_{D}*0 + \beta_{E}*0 + \epsilon$$ 
$$y = \beta_0 + \beta_{B}+ \epsilon$$ 
etc.


Run an lm and use car::Anova(yourlmmod). How do the estimates and statistical tests we've discussed relate to the visual representation in your figures. 


# Apply to all sets

Load the other data and run this plotting function for each of the datasets as well as your model.
How does sample size impact your conclusions? How does the variance in the error impact them?

